---
title: "Dimensionality Reduction & Clustering"
author: "Kaitlin Sullivan"
format: html
editor: visual
---

```{r, include=FALSE}
library(RUHi)
myobj <- readRDS("demo2.rds")
```

::: {style="display: flex; justify-content: space-between; margin-top: 2em;"}
<a href="06_filtering-processing.qmd" class="btn btn-outline-secondary">⟵ Previous: Filtering & Processing</a> <a href="08_visualization.qmd" class="btn btn-primary">Next: Visualization ⟶</a>
:::

## Video Tutorial

The following video tutorial demonstrates the functionality of `ruUMAP`() and `ruCluster()`, two functions that dimensionally reduce and hierarchically (method = "ward.D2") cluster the data.

*Note that UMAPs have built-in stochasticity, meaning they will look different between computers*

Follow along with the code below.

::: {.center}
<iframe src="https://www.youtube.com/embed/BZINf1FQkec" width="640" height="360" allowfullscreen></iframe>
:::

## Running a UMAP with  `ruUMAP()`

Once run, you can access the UMAP coordinates in the attributes (`@attributes$umap`).

```{r}
#populate attributes with UMAP for plotting
#you can use the optional argument npc to  select more or less PCs
myobj <- ruUMAP(myobj)
```

## Clustering with  `ruCluster()`

This function clusters the data using ward D2 hierarchical clustering. Select a number of clusters you beleive the data might have. Once run, a new `@metaData` column will be generated. 

To refine the number of clusters, check the dendrogram with `plotDendro()` and re-run `ruCluster()` as many times as you see fit.

```{r}
#populate metaData with cluster column
myobj <- ruCluster(myobj, k = 5)

#check the dendrogram to see if you agree with number of clusters
plotDendro(myobj)
```


```{r include=FALSE}
saveRDS(myobj, "demo3.rds")
```


::: {style="display: flex; justify-content: space-between; margin-top: 2em;"}
<a href="06_filtering-processing.qmd" class="btn btn-outline-secondary">⟵ Previous: Filtering & Processing</a> <a href="08_visualization.qmd" class="btn btn-primary">Next: Visualization ⟶</a>
:::
