---
title: "Visualization"
author: "Kaitlin Sullivan"
format: html
editor: visual
---

```{r, include=FALSE}
library(RUHi)
myobj <- readRDS("demo3.rds")
```

::: {style="display: flex; justify-content: space-between; margin-top: 2em;"}
<a href="07_dimred-clustering.qmd" class="btn btn-outline-secondary">⟵ Previous: Dimensionality Reduction & Clustering</a> <a href="09_geographic-analysis.qmd" class="btn btn-primary">Next: Geographic Analysis ⟶</a>
:::

## Video Tutorial

The following video tutorial demonstrates some of the ways you can visualize mFISH data.

See code below for some basic plotting functions.

::: center
<iframe src="https://www.youtube.com/embed/DKMQuPOA8Ns" width="640" height="360" allowfullscreen>

</iframe>
:::

## Plotting Functions

#### Geographic Space

```{r}
######### STEP 4: PLOTTING #########

### GEOGRAPHIC SPACE with plotSpace()
#plot in space - automatically coloured by cluster
plotSpace(myobj)

#optional args to group by section, or other variable (eg cluster)
plotSpace(myobj, group.by = 'cluster')

#Plot by metadata values
library(dplyr)
myobj@metaData <- mutate(myobj@metaData, virus=ifelse(LEC>3, 'LEC', ifelse(RSC>3, 'RSC', 'None')))
plotSpace(myobj, colour.by = "virus")

#plot in space but change to a gene or metadata value
plotSpace(myobj, colour.by = 'Gnb4', include.fil = F)

#plot in space with separation by cluster (group.by is useful for viewing multiple sections as well)
plotSpace(myobj, group.by = 'cluster', colour.by = 'Gnb4')
```

#### Dimensionally Reduced Space

```{r}
### DIM REDUCED SPACE with plotDim()
#auto coloured by cluster
plotDim(myobj)

#option to colour by gene/metadata 
plotDim(myobj, colour.by='Gnb4')
plotDim(myobj, colour.by = 'virus')
```

#### Gene Expression Box Plots

```{r}
### MARKER GENE BOX PLOTS

#Plot a gene's expression across clusters
geneBoxPlot(myobj, 'Gnb4')

#Plot the gene expression profile of a specified cluster
clusterBoxPlot(myobj, clus='5')

#or simply plot the gene expression for every cluster
clusterBoxPlot(myobj)
```

::: {style="display: flex; justify-content: space-between; margin-top: 2em;"}
<a href="07_dimred-clustering.qmd" class="btn btn-outline-secondary">⟵ Previous: Dimensionality Reduction & Clustering</a> <a href="09_geographic-analysis.qmd" class="btn btn-primary">Next: Geographic Analysis ⟶</a>
:::
